---
title: "SQL Exercise"
output: html_document
date: "2024-10-15"
---


First create a database connection in R: 

```{r}
database <- dbConnect(RSQLite::SQLite(), "C:/Users/neilb/navco_db.db") 

dbListTables(database)
```

Use this code to view the first few rows

```{sql}
#| connection = database

SELECT * FROM NAVCO_12 LIMIT 10

```


If you want to be able to access the results of a query in R, add the output.var option to the sql chunk: 

```{sql, output.var = tenrows}
#| connection = database

SELECT * FROM NAVCO_12 LIMIT 10

```

Alternatively, you can just use the `dbGetQuery` function from R: 

```{r}

dbGetQuery(database, "SELECT * FROM NAVCO_12 LIMIT 10")

```

# Question 1

Get the total number of rows in NAVCO 1.2

```{sql}


```

# Question 2

Calculate the average %  participation for violent and non-violent movements. Put your result in an R variable and then create a barplot showing your results

```{sql}


```


# Question 3

Do any countries have more than one maximalist campaign? Which had the highest?

```{sql}


```

# Question 4

Use SQL or R to produce a count of each type of campaign


```{sql}


```



