---
title: "SQL Exercise"
output: 
  html_document:
      df_print: paged
      code_download: true
      toc: true
      toc_depth: 3
      toc_float: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning= FALSE, message=FALSE)

```



First create a database connection in R: 

```{r}
library(tidyverse)
library(DBI)

database <- dbConnect(RSQLite::SQLite(), "C:/Users/neilb/Documents/APAN/GVPT628-Fall2024/NAVCO.db") 

dbListTables(database)
```

Use this code to view the first few rows

```{sql}
#| connection = database

SELECT * FROM NAVCO_V2 LIMIT 10

```


If you want to be able to access the results of a query in R, add the output.var option to the sql chunk: 

```{sql, output.var = 'tenrows'}
#| connection = database

SELECT * FROM NAVCO_V2 LIMIT 10

```

Alternatively, you can just use the `dbGetQuery` function from R: 

```{r}

dbGetQuery(database, "SELECT * FROM NAVCO_V2 LIMIT 10")

```

# Question 1

Get the total number of rows in NAVCO V2



# Question 2

Calculate the average %  participation for violent and non-violent movements. Put your result in an R variable and then create a barplot showing your results



# Question 3

Do any countries have more than one maximalist campaign? Which had the highest?


# Question 4

Use SQL or R to produce a count of each type of campaign




