---
title: "ggplotting"
output: 
  html_document:
      df_print: paged
      code_download: true
      toc: true
      toc_depth: 3
      toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE)
```

Load in data on races from FiveThirtyEight

```{r, warning=FALSE}
library(tidyverse)
url <- 'https://raw.githubusercontent.com/fivethirtyeight/election-results/refs/heads/main/races.csv'
races<-read_csv(url)
head(races)


```

# Q1

How would I convert these dates to a proper R format?

```{r}

races|>select(date)|>head()

```

# Q1B

This same issue shows up in all of the FiveThirtyEight election data. How would we write a function to automatically make this correction for this data?

```{r}


```

Take a look at the most recent elections (use `slice_max`). What's wrong here?

```{r}




```

# Q2

What is the most recent race to have occurred? (note that some of these haven't happened yet)

# Q3

These results appear to be updated sporadically. So how could I find the most recent race or races for which outcome data are available?

```{r}





```

# Q4

How would I get all races in the next 120 days regardless of when I run the code? In other words, write something that extracts upcoming races relative to `Sys.Date()`?

```{r}



```

# Q5

Access the House Race data from the [FiveThirtyEight github](https://github.com/fivethirtyeight/election-results/tree/main). What race has the narrowest margin of victory in the last 10 years?

```{r}


```

# Q5B

Create a plot showing your results

```{r}


```
