---
title: "Webscraping Exercise"
output: 
  html_document:
      df_print: paged
      code_download: true
      toc: true
      toc_depth: 3
      toc_float: true
---

```{r setup, include=FALSE}


knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

```{r}
library(tidyverse)
library(rvest)

```

Goal: we're going to create a data frame with all of the current travel advisories on the U.S. State Department website, along with their full text and the relevant advisory tags. We'll use the `rvest` package to accomplish this. 

Start by reading in the data

```{r}

url<-'https://travel.state.gov/content/travel/en/traveladvisories/traveladvisories.html/'

result<-read_html(url)

```

The travel advisories are in an HTML table object. Fortunately, `rvest` offers a simple function to extract and re-format an HTML table: `html_table()`

```{r}
url<-'https://travel.state.gov/content/travel/en/traveladvisories/traveladvisories.html/'

travel_advisories<-result|>
  html_element('table')|>
  html_table()

```

# Question 1

We have the table object but how do we extract the hyperlinks? Use the Selector Gadget to get a css selector statement and extract the links to each travel alert.

```{r}
# Q1 code


```



We have relative URLs, but remember that we can complete a relative URL with the function below:

```{r, eval=FALSE}
url<-'https://travel.state.gov/content/travel/en/traveladvisories/traveladvisories.html/'

abs_path<-xml2::url_absolute(links, url)

travel_advisories$links<-abs_path

```

Now we have a list of links. These will all have similar formatting, so if you figure out how to scrape one you can probably grab the rest without much problem.

## Question 2

Visit the first link and write a selector statement to extract the alert text and the threat tags and assign them to objects in R.

```{r}
# Q2 code


```

## Question 3

Now: you already have a data frame that can store this object, so you just need to create two empty columns and then create a loop that will visit each link, extract the text and advisory tags, and place them in the appropriate cell. Be sure to put a short pause between each iteration of the loop.

```{r}
travel_advisories$text<-NA
travel_advisories$tags<-NA
```

Hint: Since there can be multiple "threat" tags for a single advisory, you can create a list column in your loop like this:

```{r, eval=FALSE}
# inside the loop: 
travel_advisories$tags[i] <- list(threats)
```

```{r}
# Q3 code

```

