---
title: "API Queries"
output: 
  html_document:
      df_print: paged
      code_download: true
      toc: true
      toc_depth: 3
      toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
current_congress_number<- 118
```

Our goal in this analysis is to get the age information about all members of the current congressional delegation. We can use the member_details endpoint to get detailed information about a member, but we need to know a members `bioguideId` in order to get their year of birth, so we're actually going to need to send queries to multiple different endpoints in order to assemble this data all in one place.

Sign up for an API key here: <https://api.congress.gov/sign-up/>

Query documentation: <https://api.congress.gov/>

Detailed data documentation: <https://github.com/LibraryOfCongress/api.congress.gov/tree/main/Documentation>

Start by loading the relevant libraries and creating a base URL that we can re-use. Since we'll always need to send our API key, you can go ahead and add this as a parameter to the base URL object:

```{r}
library(httr2)
library(tidyverse)
# get the base URL 
request = request('https://api.congress.gov/v3/')|>
  # append the API key 
  req_url_query('api_key'= Sys.getenv("CONGRESS_API_KEY"))

```

We can use this code to retrieve information about the current congress:

```{r}

current_congress<-request|>
  req_url_path_append("congress/current")|>
  req_perform()|>
  resp_body_json()


```

## Question 1

Examine the structure of `current_congress` response and extract the number of the current congress (it should be just a number) and assign it to an object called `current_congress_number`

```{r}



```

According to the [API documentation](https://api.congress.gov/) we can extract member information from the `member` endpoint. The query should look something like this: \
https://api.congress.gov/v3/member/congress/118?api_key=[INSERT_KEY]

To make our code more future-proof, we can replace `118` with the `current_congress_number` variable that we created in the prior step.

```{r}

congress_info<-request|>
  req_url_path_append("member","congress", current_congress_number)|>
  req_url_query(limit =250)|>
  req_perform()|>
  resp_body_json()


# take a look at the first object
congress_info$members[[1]]
congress_info$pagination

```

This only gives us the first 250 members, lets go ahead and use a loop to extract data on all 535 members.

```{r, cache=TRUE}



congress_list <- congress_info$members
# then create our list of offsets
offsets<-seq(from=250, to = 552, by= 250)

# iterate over the offsets and add them as the i-th element of the list
for(i in offsets){
  next_page <-
    congress_info <- request |>
    req_url_path_append("member", "congress", current_congress_number) |>
    req_url_query(limit = 250, offset=i) |>
    req_perform() |>
    resp_body_json()
  
  congress_list <- c(congress_list, next_page$members)
  
}


```

We can extract some specific member information like this:

```{r}


variables<-c('name', 'state', 'partyName', 'url')

congress_list[[1]][variables]



```

## Question 2

Write a loop to iterate through values of `congress_list` and populate the empty data frame below:

```{r}

nrows <- length(congress_list)

congress_df <- data.frame('name'=character(nrows),
                          'state' = character(nrows),
                          'partyName'=character(nrows),
                          'url'  = character(nrows)
                          )




```



## Question 3A

Finally, we can use the `url` columns (along with our API key) to get detailed biographical information about a member.

Write a loop that iterates through each value of `url` in `congress_df` and queries the relevant member information (you can just do the first 20 for the sake of keeping things to a reasonable time frame)

```{r}
#

```

## Question 3B

Now, extract each member's birth year and then use it to calculate age of the current congressional delegation. Change it to numeric and create a density plot or boxplot to show the distribution of congressional ages.

```{r}
#


```
