---
title: "Homework 2 Solutions"
author: Neil Lund
output: 
  html_document:
      df_print: paged
      code_download: true
      toc: true
      toc_depth: 3
      toc_float: true
---

```{r setup, include=FALSE, cache=T}

# good default settings for homeworks (suppress warning messages but always show code and output)
knitr::opts_chunk$set(echo = TRUE, 
                      
                      warning =FALSE,
                      message=FALSE)

# if you have code that runs for a long time, add cache=TRUE to this set of options, this will prevent the code from being re-evaluated every time you knit the document


```

# Preamble

Loading the census data on county incomes. (You weren't required to use this, but it can be useful for an example)

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(tidycensus)

county_incomes <- get_acs(geography = "county", 
              variables = c(medincome = "B19013_001"), 
              year = 2020)|>
  filter(!is.na(estimate))



```

# Q1

We already have the loop for getting a bootstrapped confidence interval for a median. We just need to repeatedly resample the data with replacement and then get the 2.5th and 97.5th quantiles of the results. The basic loop looks like this:

```{r}

# Set the random number seed 
set.seed(99) 


# create an empty vector to store results
sampled_medians<-vector("numeric", length=3000) 


for(i in 1:3000){
  # sample with replacement
  sample_n <- sample(county_incomes$estimate, 
                     size=nrow(county_incomes), 
                     replace=T)
  
  # now calculate our statistic of interest for this new sample:
  stat<-median(sample_n)
  
  # and add the result to the empty list of sample medians
  sampled_medians[i] <- stat
  
}


```

Now we can use the sampled medians to get a confidence interval. (we can also use the 50th percentile as our point estimate if we want, instead of using the actual sample median)

```{r}



quantile(sampled_medians, c(0.025,.5,.975))

```

To generalize this for a function, we basically just need to swap out the part that is specific to our dataset `county_incomes$estimate` for something more generic that we can supply when we call the function. In addition, I've replaced `for(i in 1:3000)` with `for i in 1:nsamples` and made `nsamples` one of the arguments to my function. That way I can also change the number of new samples I generate when I run the boot strap. (In general, a few hundred is fine, but more is always better provided you have the time)

```{r}


bootMedian<-function(x, nsamples=200){
  sampled_medians<-vector("numeric", length=nsamples) # empty numeric vector of length n
  for(i in 1:nsamples){
    # sample with replacement
    sample_n <- sample(x, 
                       size=length(x), 
                       replace=T)
    # now calculate our statistic of interest for this new sample:
    stat<-median(sample_n)
    
    # and add the result to the empty list of sample medians
    sampled_medians[i] <- stat
  }

  result<-c(mean(sampled_medians), quantile(sampled_medians, c(0.025, .975)))
  
  names(result) <- c("mean", "lower ci", "upper ci")
  return(result)

}

bootMedian(county_incomes$estimate)


```

# Q1 Bonus

We can add a lot of additional options to this basic function.

The `level` argument lets me customize the size of my confidence interval.

The `allsamples` argument allows us to optionally return the full list of bootstrapped replicates, instead of just the quantiles (note that I use a list to return multiple objects from the function).

The `seed` argument allows me to control the random number seed when I call the function. (remember that the random number seed controls how random data gets generated by R, so its a way of ensuring that we can replicate an analysis that relies on a random process)

Lastly: instead of using the mean as the point estimate from the bootstrapping, I'll use the 50th percentile. (this wasn't required, but its a valid option that could be more sensible in this case)

```{r}
bootMedianFancy<-function(x, n=200, 
                          level=.95, # sets the confidence level (default =95%)
                          allsamples=FALSE, 
                          seed=NULL # allows you to set a random number seed
                          ){
  # if user supplies a random number seed, then set it here: 
  if(!is.null(seed)){
    set.seed(seed)
  }
  # get the vector of quantiles you'll be extracting 
  ci_quantiles<-c((1 - level)/2,  # lower bound ci
             .5,   # the median 
             1-(1 - level)/2 # upper bound ci
             ) 
  sampled_medians<-vector("numeric", length=n) # empty numeric vector of length n
  
  for(i in 1:n){
    # sample with replacement
    sample_n <- sample(x, 
                       size=length(x), 
                       replace=T)
    
    # now calculate our statistic of interest for this new sample:
    stat<-median(sample_n)
    
    # and add the result to the empty list of sample medians
    sampled_medians[i] <- stat
  }
  result<-list()
  result$CI<-quantile(sampled_medians, ci_quantiles)
  
  # if the user wants all the samples, then return them along with
  # the CI estimate
  if(allsamples == TRUE){
    result$samples <- sampled_medians
  }
  return(result)
}



booted<-bootMedianFancy(county_incomes$estimate,  
                        level=.90,  # set a 90% CI instead of 95%
                        allsamples=TRUE, # return all the samples
                        seed=100 # set the random number seed to 100
                        )

booted$CI



```

## Extending it further

This is not one of the suggested options, but we could actually make this function even more generalizable by adding a `fun` argument that allows us to a pass another function to our function. Instead of just getting the confidence interval for a median, we can generate confidence intervals for pretty much any statistic that takes a series of numbers and returns a single value.

```{r}
bootAnything<-function(x, fun, n=200, level=.95, allsamples=FALSE, seed=NULL){
  
  # if user supplies a random number seed, then set it here: 
  if(!is.null(seed)){
    set.seed(seed)
  }
  # get the vector of quantiles you'll be extracting 
  ci_quantiles<-c((1 - level)/2,  # lower bound ci
                  .5,   # the median 
                  1-(1 - level)/2 # upper bound ci
  ) 
  sampled_results<-vector("numeric", length=n) # empty numeric vector of length n
  
  for(i in 1:n){
    # sample with replacement
    sample_n <- sample(x, 
                       size=length(x), 
                       replace=T)
    
    # now calculate our statistic of interest for this new sample:
    stat<-fun(sample_n) # note: "fun" can be any function that takes a vector and 
    # returns a single value
    
    # and add the result to the empty list of sample medians
    sampled_results[i] <- stat
    
  }
  
  result<-list()
  result$CI<-quantile(sampled_results, ci_quantiles)
  
  # if the user wants all the samples, then return them along with
  # the CI estimate
  if(allsamples == TRUE){
    result$samples <- sampled_results
  }

  return(result)
}



```

Here's an example of using the function above to get confidence intervals for the mean absolute deviation (an alternative to the standard error that is more robust to outliers and skewness).

```{r}

mad <- function(x){
  return(sum(abs(x - mean(x)))/length(x))
  
  }

bootAnything(county_incomes$estimate, 
             fun=mad,
             seed=100)



```

# Q2

For this question, we can generate some example data by using the sample function and R's built in list of the letters of the alphabet. In this case we can tell the modal value is "k" just by using `sort` with `table`, but we want to create a function for this.

```{r}

set.seed(9999)
sampledata<-sample(letters, size=1000 ,replace=T)

sort(table(sampledata), decreasing=T)

```

While its probably fine to call this function `mode`, there's actually already a base R function with that name that does something else. So to avoid confusion we'll call our function `modalValue`. It should take a vector and should throw an error if the vector is not a character type.

The return value here should be a letter. So we should return the name of the most frequent item in the table, instead of the count:

```{r}

modalValue<-function(x){
  if(class(x)!='character'){
    stop("x is not a character vector")
  }
  
  m <- sort(table(x),decreasing=T)[1]
  
  
  
  
  return(names(m))
}



modalValue(sampledata)



```

(as an aside, requiring the vector to be character is kind of arbitrary. Modes are equally valid for numeric variables or R `factor` types I just wanted you to mess with using control flow and error messages)

## Making it faster

There's [a thread on stackoverflow](https://stackoverflow.com/questions/2547402/how-to-find-the-statistical-mode) that has a number of different suggestions for getting the statistical mode of a vector. These will all give you the same result, but looking at different ways to solve the same problem can be useful way to learn more about a language.

The most upvoted answer from user Ken Williams is below. It will give us the same result, but you might note that its significantly faster than what I wrote above. Not that this is a slow process in either case, but looking at the code for the `tabulate` and `table` command might offer some insights as to why there's a noticeable speed boost. (you can view the code for a function in the console by referencing the function name without parentheses)

```{r, warning=FALSE}
library(microbenchmark) # a package for benchmarking code

Modes <- function(x) {
  ux <- unique(x)
  tab <- tabulate(match(x, ux))
  ux[tab == max(tab)]
}

set.seed(9999)
sampledata<-sample(letters, size=10000 ,replace=T)

comparison<-microbenchmark( Modes(sampledata), modalValue(sampledata))


# View the results of the speed comparison: 
summary(comparison)

```
