---
title: "Python from R"
output: 
  html_document:
      df_print: paged
      code_download: true
      toc: true
      toc_depth: 3
      toc_float: true
---


<!--When using Python in Rmarkdown, add: library(reticulate) use_virtualenv(...)
to the setup chunk. This ensures that you'll get the right versions and packages
and whatnot.-->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
library(reticulate)
use_virtualenv("APAN-Python")


```

To run python code, just replace the `r` part of the chunk with `python`.

```{python}

print("python code")


```

You can still use R code in the same document:

```{r}
#install.packages("httr2")
library(httr2)
library(tidyverse)
battle_deaths<-request('https://ucdpapi.pcr.uu.se/api/battledeaths/23.1?pagesize=10&Country=365')|>
  req_perform()|>
  resp_body_json()


result<-bind_rows(battle_deaths$Result)




```

# Comparing R and Python

Python, like R, is an interpreted language that can be used interactively. A lot of the same caveats about loops and memory usage will apply to Python just as much as they do to R.

In terms of syntax, Python code can look a lot like R code. For instance: using functions will follow the same `function(arguments)` type of syntax that we've seen with R. (you might also notice that we can add comments the same way)

```{python}
# printing hello
print("hello")


```

Listing all of the differences between R and Python syntax isn't really feasible, but there are some differences that will pop out immediately that we can discuss here.

## 1. Indexing

Python allows you to use the `[]` to get a subset of some data, but Python uses 0-based indexing, so the first item in a list is the 0th element.

```{python}
mylist = ["a", "b", "c", "d"]

mylist[0]

mylist[1]

```

We can also access multiple elements using a range like `[0:3]`, but the range will stop before the second number, so `[0:4]` means elements 0,1,2,3 instead of 0,1,2,3,4:

```{python}
mylist[0:3]

```

## 2. Loops, control flow and functions

Python uses indentation to do a lot of the stuff you would do with `{}` in R. For instance, here's how we would loop through an R list:

```{r}

vec<-c("fee", "fi", "fo","fum")

for(i in vec){
  print(i)
}

```

And here's how we would do it in Python:

```{python}
vec = ["fee", "fi", "fo", "fum"]

for i in vec:
  print(i)

```

Here's how we would create a function in R

```{r}
numvec<-c(1, 1.5, 10, 11, -1,13)

se<-function(x){
  x<-sd(x)/sqrt(length(x))
  return(x)
}

se(numvec)

```

And here's an equivalent in Python:

```{python}


numvec = [1, 1.5, 10, 11, -1,13]


def se(x):
  mu = sum(x)/len(x)
  sigma = (sum([(i-mu)**2 for i in x])/len(x)) **.5
  s = sigma/((len(x)-1)**.5)
  return(s)

se(numvec)

```

## 3. Methods and classes

Methods and classes also exist in R, but they're a little more obscured. A method is a function that is associated with a particular class. So the `.lower` method is a function that only works on variables that have the string class:

```{python}

# this works
words = 'SOME WORDS'
words.lower()

```

```{python, error=TRUE}
# but this gives an error message
numbers = 1

numbers.lower()
```

You'll often see method functions called using this `objectname.functionname()` syntax in Python.

## 4. Modules

Python packages work very similarly to R packages with a couple of minor differences.

First, package installation will usually be done outside of Python using the pip command. You can do this by opening a terminal window and typing in `pip install [packagename]` or you can do it from within R by opening a terminal tab and typing the same.

Alternatively, you can use the `virtualenv_install` function from the reticulate package to install a package in a virtual Python environment, or use `py_install` to install the package in your native Python installation

```{r, eval=FALSE}
library(reticulate)

virtualenv_install(envname="APAN-Python", "pandas")




```

To use a function from a Python module, you'll usually need to prefix the module name followed by a `.` and then the function name. ie. `numpy.std()` or `pandas.mean()`. To make this a little less verbose, you can give an alias to the package by importing like `import [package] as [alias]`, then you can use that alias in place of the full name of the package:

```{python}
import requests as req

battledeaths = req.get('https://ucdpapi.pcr.uu.se/api/battledeaths/23.1?pagesize=10&Country=365')

bd_json = battledeaths.json()


```

## 5. General design philosophy

Python is more "programmer oriented" where R is more oriented towards statisticians. So things that seem simple in one language can be convoluted in another. You can see an example of this where we calculated the standard error above: we had to write a lot more code because there's no native Python function for calculating something as simple as mean or standard deviation. Fortunately, we can usually overcome this with packages like `numpy` but it reflects a difference in the user base of the languages. 

One package that you'll see a lot in Python data analysis is `pandas`. It provides a lot of the same functionality that we get from the `tidyverse` packages in R, including a lot of support for easy split-apply-combine functions. 


```{python}

import pandas as pd

conflict_data = pd.json_normalize(bd_json["Result"])


conflict_data.head()

```



## Working between R and Python



The `reticulate` package allows you to reference objects in either the R or Python environment, so you can pass results back and forth between the languages.

I can create this in R:

```{r}
set.seed(100)
rvector = rnorm(10)

```

And then I can access it in a Python chunk by writing `r.[object name]`:

```{python}

pvec = sum(r.rvector)

pvec

```

Then I can grab the object back from Python by running `py$[object name]`:

```{r}

py$pvec

```

This kind of workflow might not be ideal for huge data sets, but it works fairly well for the kinds of quick trips to Python that we might make in this course.
