---
title: "Homework 1 Solutions"
author: Neil Lund
output: 
  html_document:
      df_print: kable
      code_download: true
      toc: true
      toc_depth: 3
      toc_float: true
---

```{r setup, include=FALSE}

# good default settings for homeworks (suppress warning messages but always show code and output)
knitr::opts_chunk$set(echo = TRUE, 
                      warning =FALSE,
                      message=FALSE)

# if you have code that runs for a long time, add cache=TRUE to this set of options, this will prevent the code from being re-evaluated every time you knit the document


```

# Preamble

```{r}
library(tidyverse)

# download the CHES special edition data:
ches<-haven::read_dta('https://www.chesdata.eu/s/CHES_Ukraine_March_2024.dta')

# convert stata labelled vectors to R factors
ches<-as_factor(ches)

```

## Q1

**What is the most common party family among the 30 countries studied?**

A. The liberal party family

There's a lot of ways to do this, one of the easiest is probably `table`+ `sort`

```{r}

table(ches$family)|>
  sort(decreasing=T)


```

An alternative method is to use the dplyr function `count` with `arrange`. This returns a `tibble` instead of a named vector, which can be easier to use (and will look nicer in markdown if we set `df_print: kable` in the markdown options)

```{r}
ches|>
  count(family, sort=TRUE)

```

## Q2

**What is the most successful radical left party based on current seat share? (give the party name and country)**

Syriza (Greece) with 23.6% of seats.

A base R method could use `order` or `which.max`

```{r}
radleft <- ches[which(ches$family=='rad left'),]

# get the index of the highest value using order:
maxvalue<-order(radleft$seatperc, decreasing=T)[1]

# or use which.max()
maxvalue<-which.max(radleft$seatperc)

radleft[maxvalue,c('party', 'country', 'seatperc')]



```

An equivalent using the `dplyr` package:

```{r}
ches|>
  select(party, country, seatperc)|> # get just the relevant cols
  arrange(-seatperc)|> # sort from highest to lowest
  slice_head(n=1) # take the first row

```

## Q3

**Calculate the average and the standard deviation of UA_EU (average support for Ukraine joining the European Union) for each party family. What family has the highest average? What family has the highest variance in their level of support?**

Liberal have the highest average. Traditional/Authoritarian/Nationalist parties have the highest variance

You can use aggregate:

```{r}

# aggregate(ches$UA_EU, list(ches$family) , FUN=mean)

# OR: 

aggregate(data=ches, UA_EU ~ family, FUN=mean)
aggregate(data=ches, UA_EU ~ family, FUN=sd)

          
```

You can also use `dplyr`'s `summarize` with `group_by`, which allows you to create a single data frame with both metrics included

```{r}

ches|>
  group_by(family)|>
  summarise(mean_sup = mean(UA_EU),
            sd_sup = sd(UA_EU)
            )

```

## Q4 

**Calculate the effective number of political parties. for a single country of your choosing based on either seat-share or vote-share**

The ENPP measure is:

$$N = \frac{1}{\sum^n_{i=1} p_i^2} $$ Where `p` is either a seat share or a vote share.

We can calculate this for Hungary:

```{r}
# get the rows for Hungary:
hungary_rows<-ches$seatperc[which(ches$country=="Hungary")]

# normalize so they sum to 1
normalized<-hungary_rows/sum(hungary_rows)

# now calculate:
1/sum(normalized^2)

```

## Q5

**Write a function that takes the number of seats held by each party in a given country and returns the effective number of political parties.**

The simplest version more-or-less just generalizes from what we used above. Replacing the specific seatshares vector with "x":

```{r}
enp<-function(x){
  x<-x/sum(x) # normalize shares
  result<-1/sum(x^2)
  return(result)
  
}




```

Now we can use this function with aggregate to get results for all countries at once:

```{r}

aggregate(ches, seatperc ~ country, enp)

```

Or use `dplyr` functions to get electoral parties based on vote share, and parliamentary parties based on seats:

```{r}

ches|>
  group_by(country)|>
  summarise(enpp = enp(seatperc),
            enep = enp(vote)
            )

```



